Holomorphic field theories and higher algebra

### 2023-07-27 highervertexops.tex
short note

The goal here is to record higher-dimensional analogs of the ingredients for defining a {\em vertex operator}. We do not aspire (yet!) to give the higher-dimensional version of a vertex algebra.

Let $\cO_d$ denote a (preferably ``small'' or convenient) model of the derived sections of the structure sheaf (of {\em algebraic} functions) on the punctured affine space $\mathring{\AA}^d$.
When $d=1$, it is the Laurent polynomials~$\CC[z,z^{-1}]$.

Let $\widehat{\cO}_d$ denote a (preferably ``small'' or convenient) model of the derived sections of the structure sheaf (of {\em algebraic} functions) on the punctured formal disk $\mathring{\DD}^d$.
When $d=1$, it is the Laurent series $\CC((z))$.

Let $\cD_d$ denote a model of the derived ``distributions'' on the punctured affine space $\mathring{\AA}^d$. It is the linear dual to~$\cO_d$.
When $d=1$, it is $\CC[[z,z^{-1}]]$, which is a notation that indicates infinite formal sums in both directions. (This notation is poor because it suggests an algebra, but it is not.)
The pairing is via the residue pairing.

Let $\widehat{\cD}_d$ denote a model of the derived ``distributions'' on the punctured formal disk $\mathring{\DD}^d$. It is the linear dual to~$\widehat{\cO}_d$.
When $d=1$, it is again $\CC((z))$, via the residue pairing.

@@

A {\em field} is an element of 
\[
\RHom(V \otimes \widehat{\cO}_d, V) \cong \RHom(V, V \widehat{\otimes} \widehat{\cD}_d).
\]
When $d=1$, this becomes the usual version $\Hom(V, V((z)))$.

@@

Given two fields $a(z)$ and $b(w)$, we have a commutator $[a(z),b(w)]$ that eats a

>>>>

The goal here is to record higher-dimensional analogs of the ingredients for defining a {\em field} in a holomorphic theory (i.e., a higher-dimensional {\em vertex operator}). We do not aspire (yet!) to give the higher-dimensional version of a vertex algebra.

{\it Notations for spaces, functions, and distributions}

Let $\cO_d$ denote a (preferably ``small'' or convenient) model of the derived sections of the structure sheaf (of {\em algebraic} functions) on the punctured affine space $\mathring{\AA}^d$.\\
{\tiny When $d=1$, it is the Laurent polynomials~$\CC[z,z^{-1}]$.}

Let $\widehat{\cO}_d$ denote a (preferably ``small'' or convenient) model of the derived sections of the structure sheaf (of {\em algebraic} functions) on the punctured formal disk $\mathring{\DD}^d$.\\
{\tiny When $d=1$, it is the Laurent series $\CC((z))$.}

Let $\cD_d$ denote a model of the derived ``distributions'' on the punctured affine space $\mathring{\AA}^d$. It is the linear dual to~$\cO_d$.\\
{\tiny When $d=1$, it is $\CC[[z,z^{-1}]]$, which is a notation that indicates infinite formal sums in both directions. The pairing is via the residue pairing. (This notation is poor because it suggests an algebra, but it is not.)}

Let $\widehat{\cD}_d$ denote a model of the derived ``distributions'' on the punctured formal disk $\mathring{\DD}^d$. It is the linear dual to~$\widehat{\cO}_d$.\\
{\tiny When $d=1$, it is again $\CC((z))$, via the residue pairing.}

There are analogs of these constructions for products of affine or formal spaces with the fat diagonal removed.
We will write here the two-variable version.

Consider the space 
\[
\mathring{\AA}^{d}[2]= (\AA^d \times \AA^d) - \Delta_f{[2]}, 
\]
where $\Delta[2]_f$ denotes the fat diagonal
\[
\AA^d \times \{0\} \cup \{0\} \times \AA^d \cup \AA^{2d}_\Delta
\]
with
\[
 \AA^{2d}_\Delta = \{ (x,y) \in \AA^d \times \AA^d \; : \; x \neq y\}.
\]
Let $\mathring{\DD}^d[2]$ denote the corresponding formal space.

Let $\cO_d{[2]}$ denote a model of the derived sections of the structure sheaf (of {\em algebraic} functions) on $\mathring{\AA}^d[2]$.

Let $\widehat{\cO}_d{[2]}$ denote a model of the derived sections of the structure sheaf (of {\em algebraic} functions) on $\mathring{\DD}^d[2]$.

Let $\cD_d[2]$ denote a model of the derived distributions on $\mathring{\AA}^d[2]$.

Let $\widehat{\cD}_d[2]$ denote a model of the derived distributions on $\mathring{\DD}^d[2]$.
\newpage

{\it Operators}

@@

A {\em field} is an element of 
\[
\RHom(V \otimes \widehat{\cO}_d, V) \cong \RHom(V, V \widehat{\otimes} \widehat{\cD}_d).
\]
When $d=1$, this definition becomes the usual version: $\Hom(V, V((z)))$.

@@

Given two fields $a(z)$ and $b(w)$, we have a commutator $[a(z),b(w)]$ that eats a function in $\widehat{\cO}_d[2]$ and returns an endomorphism of~$V$.

Two fields are {\em mutually local} if there exists a nonnegative exponent $N \in \NN^d$ such that $(z-w)^N[a(z),b(w)]$ is cohomological to zero (i.e., it is exact).

\owen{I suspect there's something better to say here, informed by your work with Nik. I'm least happy with this definition.}

{\it Free field realizations}

We want to show that the $L_\infty$-action of a higher Kac-Moody algebra on the state space of a free theory is via mutually local fields.
